Dual-Mode PSK Transceiver on SDR With FPGA

### 2024-01-03 latex/paper.tex
Working on Paper

\documentclass[journal,twoside]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multicol,multirow}
\usepackage{newtxmath}
\usepackage{mathtools}
\usepackage{bm}
\usepackage[caption=false,font=footnotesize,subrefformat=parens]{subfig}
\usepackage{lipsum}
\usepackage{tabularx}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{my-tikz}
\usepackage{colortbl}
\usepackage{scalerel}
\usepackage{threeparttable}
\usepackage{algorithm}
\usepackage[italicComments=false,commentColor=darkgray,noEnd=true]{algpseudocodex}
\usepackage{upgreek}
\usepackage{calc}
\usepackage[noadjust]{cite}
\usepackage{circuitikz}
\usepackage{siunitx}
\usepackage{url}
\usepackage[colorlinks,urlcolor=black,linkcolor=blue,citecolor=blue]{hyperref}

@@

  \author{%
    Wuqiong~Zhao{\hspace{.1em}\textsuperscript{\orcidicon{0000-0002-9550-7423}}},~\IEEEmembership{Student Member, IEEE}
    \thanks{Date of publication 2 January 2024; date of current version 2 January 2024.
      The author would like to thank Lecturer Dan Yang for the guidance and help in the course of the experiment.
      The author would also like to thank AI tools including GitHub Copilot and Claude.AI for their help in preparing this paper.}
    \thanks{Wuqiong Zhao is with Southeast University, Nanjing 211189, China. (e-mail: wqzhao@seu.edu.cn, website: \url{https://wqzhao.org}).}
    \thanks{Online URL: \url{https://go.wqzhao.org/sdr-psk-fpga}}
    % \vspace{-1mm}
  }

@@

      All reset signals are generated using Processor System Reset Modules \cite{xilinx:pg164},
      which can provide synchronized power-up reset signals.
      The system integrated logic analyzer (system ILA) \cite{xilinx:pg261} is used to observe the internal signals.

@@

  \section{Transmitter}

@@

  \IEEEpubidadjcol
  \section{Receiver}

@@

@@

    \textbf{Header (HDR)}.
    The header field is used to provide packet information,
    including the modulation and coding scheme (MCS) and the packet length (Length).
    The remaining bits are reserved for future use.
    The MCS field currently only determines the use of BPSK or QPSK.
    The MCSs for BPSK and QPSK are defined as `01010101' and `10101010' respectively.

@@

  \section{Experiment Results}

@@

@@

  \section{Block Diagram}

>>>>

\documentclass[journal,twoside]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multicol,multirow}
\usepackage{newtxmath}
\usepackage{mathtools}
\usepackage{bm}
\usepackage[caption=false,font=footnotesize,subrefformat=parens]{subfig}
\usepackage{lipsum}
\usepackage{tabularx}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{my-tikz}
\usepackage{colortbl}
\usepackage{scalerel}
\usepackage{threeparttable}
\usepackage{algorithm}
\usepackage[italicComments=false,commentColor=darkgray,noEnd=true]{algpseudocodex}
\usepackage{upgreek}
\usepackage{calc}
\usepackage[noadjust]{cite}
\usepackage{circuitikz}
\usepackage{siunitx}
\usepackage{minted}
\usepackage[colorlinks,urlcolor=black,linkcolor=blue,citecolor=blue]{hyperref}

@@

  \author{%
    Wuqiong~Zhao{\hspace{.1em}\textsuperscript{\orcidicon{0000-0002-9550-7423}}},~\IEEEmembership{Student Member, IEEE}
    \thanks{Date of publication 4 January 2024; date of current version 4 January 2024.
      The author would like to thank Lecturer Dan Yang for the guidance and help in the course of the experiment.
      The author would also like to thank AI tools including GitHub Copilot and Claude.AI for their help in preparing this paper.}
    \thanks{Wuqiong Zhao is with Southeast University, Nanjing 211189, China. (e-mail: wqzhao@seu.edu.cn, website: \url{https://wqzhao.org}).}
    \thanks{Online URL: \url{https://go.wqzhao.org/sdr-psk-fpga}}
    % \vspace{-1mm}
  }

@@

      \textbf{Clock Generator}.
      All reset signals are generated using Processor System Reset Modules \cite{xilinx:pg164},
      which can provide synchronized power-up reset signals.

      \textbf{RF Data Converter}.
      This block is contains analog-to-digital converter (ADC) and digital-to-analog converter (DAC),
      enabled by a vendored AD9361 module.

      \textbf{Tx Signal Generator}.

      \textbf{Rx Processor}.

      \textbf{System ILA}.
      The system integrated logic analyzer (system ILA) \cite{xilinx:pg261} is used to observe the internal signals.

      \textbf{Constant Configurations}.
      Several parameters can be configured in this block.
      Most importantly, the mode control constants (\texttt{MODE\_CTRL}) are shown in Table~\ref{tab:mode_ctrl}.
      \begin{table}[htbp]
        \caption{Mode Control Constants}
        \label{tab:mode_ctrl}
        \renewcommand{\arraystretch}{1.2}
        \rowcolors{1}{white}{gray!15}
        \begin{tabularx}{\linewidth}{cYYYc}
          \toprule\tabvertspace
          \textbf{Mode} & \textbf{Localparam} & \textbf{Value} & \textbf{\ttfamily is\_bpsk} & \textbf{Packet} \\
          \tabvertspace\midrule
            BPSK & \texttt{MODE\_BPSK} & \texttt{4\textquotesingle b0001} (1) & \texttt{1\textquotesingle b1} & No \\
            QPSK & \texttt{MODE\_QPSK} & \texttt{4\textquotesingle b0010} (2) & \texttt{1\textquotesingle b0} & No \\
            Mixed & \texttt{MODE\_MIX~~} & \texttt{4\textquotesingle b0100} (4) & variable & Yes \\
          \bottomrule
        \end{tabularx}
      \end{table}

@@

    \IEEEpubidadjcol
    \section{Transmitter}

@@

      \newpage
      The Verilog code for the module \texttt{PN\_Gen} is shown below.
      \begin{minted}[frame=lines,fontsize=\scriptsize,highlightlines={11,18}]{verilog}
module PN_Gen # (parameter N = 5) (
  input      clk,
  output reg pn
);
  reg [N-1:0] PN_buf = 1; wire rst;
  generate
  if (N == 5)
    always @ (posedge clk)
      if (rst) begin PN_buf <= 5'd1; pn <= 0; end
      else begin
        PN_buf <= { PN_buf[3:0], PN_buf[4] ^ PN_buf[2] };
        pn <= PN_buf[4];
      end
  else if (N == 4)
    always @ (posedge clk)
      if (rst) begin PN_buf <= 4'd1; pn <= 0; end
      else begin
        PN_buf <= { PN_buf[2:0], PN_buf[3] + PN_buf[2] };
        pn <= PN_buf[3];
      end
  else ; // NOT implemented yet!
  endgenerate
  assign rst = !(|PN_buf);  // reset when PN_buf is all 0
endmodule
      \end{minted}

  \section{Receiver}

@@

      \begin{figure}[htbp]
        \centering
        \input{tikz/costas_loop}
        \caption{Costas loop for carrier synchronization with BPSK/QPSK support.}
      \end{figure}

      \begin{minted}[frame=lines,fontsize=\scriptsize]{verilog}
if (is_bpsk) begin // BPSK
  out_I_tdata <= in_I_tvalid ? in_I_tdata + in_Q_tdata : 0;
  out_Q_tdata <= in_Q_tvalid ? in_I_tdata - in_Q_tdata : 0;
end
else begin // QPSK
  out_I_tdata <= in_I_tvalid ? (in_Q_tdata[WIDTH-1] ?
    -in_I_tdata : in_I_tdata) >>> 6 : 0;
  out_Q_tdata <= in_Q_tvalid ? (in_I_tdata[WIDTH-1] ?
    -in_Q_tdata : in_Q_tdata) >>> 6 : 0;
end
      \end{minted}

@@

    \textbf{Header (HDR)}.
    The header field is used to provide packet information,
    including the modulation and coding scheme (MCS) and the packet length (Length) in bits.
    The remaining bits are reserved for future use.
    The MCS field currently only determines the use of BPSK or QPSK.
    The MCSs for BPSK and QPSK are defined as `01010101' and `10101010' respectively.

@@

  \section{Simulation Results}

    \subsection{Transmitter}

    \subsection{BPSK-Mode Receiver}

    \subsection{QPSK-Mode Receiver}

    \subsection{Mixed-Mode Receiver}

  \section{Experiment Results on SDR}

@@

      \textbf{Changing parameters on the fly}.
      AXI peripheral \cite{xilinx:pg127} can be used to change the parameters in the \texttt{Constant\_Config} on the fly,
      if the board allows.

@@

  \section{Block Diagrams}

### 2024-01-03 latex/tikz/transceiver_design.tex
Working on Paper

\begin{tikzpicture}[
  , thick
  , n/.style = {
    , draw
    , fill = myblued!15
    , align = center
  }
  , font = \small\sffamily
]
  \node [n] {RF Data\\Converter};
\end{tikzpicture}%

>>>>

\begin{tikzpicture}[
  , thick
  , n/.style = {
    , draw
    , fill = myblued!15
    , align = center
  }
  , font = \small\sffamily
  , lightning bolt to/.style={to path={
    let \p1=(\tikztostart), \p2=(\tikztotarget), \n1={veclen(\y2-\y1,\x2-\x1)} in
      (\p1) -- ($($(\p1)!0.6!(\p2)$)!\n1*.08!-90:(\p2)$) -- ($(\p1)!0.55!(\p2)$) --
      (\p2) -- ($($(\p1)!0.4!(\p2)$)!\n1*.08!90:(\p2)$) -- ($(\p1)!0.45!(\p2)$) -- 
      cycle (\p2)% Move to end point
    }}
  , null node/.style = {
    , inner sep = 0
    , outer sep = 0
  }
  , arr/.style = {
    , -latex
  }
  , arr r/.style = {
    , latex-
  }
  , node distance = 1mm and 5mm
]
  \node (RF) [n] {RF Data\\Converter};

  \node (Rx-p) [n, right = 9mm of RF] {Rx\\Processor};
  \node (Tx-p) [n, left = 9mm of RF] {Tx Signal\\Generator};

  \draw [arr] (Tx-p) -- (RF) node [midway, above] {\scriptsize16\,MHz};
  \draw [arr] (RF) -- (Rx-p) node [midway, above] {\scriptsize16\,MHz};
  \draw [arr] (Rx-p.east) -- ++ (6mm, 0) node [above right = 0 and -6mm] {\scriptsize1\,MHz};
  \draw [arr r] (Tx-p.west) -- ++ (-6mm, 0) node [above left = 0 and -6mm] {\scriptsize1\,MHz};

  \draw ([xshift = 1.5mm]RF.north west) -- ++ (0, 4mm) -- ++ (-1mm, 2mm) -- ++ (2mm, 0) node (Tx) [null node] {} -- ++ (-1mm, -2mm);
  \draw ([xshift = -1.5mm]RF.north east) -- ++ (0, 4mm) -- ++ (1mm, 2mm) -- ++ (-2mm, 0) node (Rx) [null node] {} -- ++ (1mm, -2mm);

  \fill [myyellow] ([xshift = 1mm, yshift = -1mm]Tx.center) to [lightning bolt to] ([xshift = -1mm]Rx);
  \node [above = of Tx, xshift = -1mm] {Tx};
  \node [above = of Rx, xshift = +1mm] {Rx};

  % \node (clk) [n, below = 4mm of RF] {Clk.\ Gen.};
  % % \draw [arr] (RF) -- (clk);
  % \draw [arr] (clk) -| ([xshift = -3mm]Rx-p.south);
  % \draw [arr] (clk) -| ([xshift = +3mm]Tx-p.south);

  % \node (cfg) [n, below = 4mm of clk] {Constant\\Config.};
  % \draw [arr] (cfg) -| ([xshift = 3mm]Rx-p.south);
  % \draw [arr] (cfg) -| ([xshift = -3mm]Tx-p.south);

  \node (clk) [n, below = 5mm of RF, xshift = -10mm] {Clk.\\Gen.\vphantom{g}};
  \node (cfg) [n, below = 5mm of RF, xshift = 7mm] {Constant\\Config.};
  \draw [arr] (clk.north) -- ++ (0, 2mm) -| ([xshift = 3mm]Tx-p.south);
  \draw [arr] (cfg.south) -- ++ (0, -2mm) -| ([xshift = 3mm]Rx-p.south);
  \draw [arr] (clk.north) -- ++ (0, 2mm) -| ([xshift = -3mm]Rx-p.south);
  \draw [arr] (cfg.south) -- ++ (0, -2mm) -| ([xshift = -3mm]Tx-p.south);

  \node [n, above = 4mm of Rx-p.north east, anchor = south east] {System ILA};

\end{tikzpicture}%
